Data for supplement to ms. 1

### 2017-04-12 ms1/ms1.Rmd
updated

summarize_by = function(bene, pde, by) {
  # e.g., by='sex'
  # count number of beneficiaries in each group
  denom = count_(bene, by) %>% rename(n_bene=n)

  pde %>%
    left_join(select_(bene, 'bene_id', by), by='bene_id') %>% # join information about the bene's sex (e.g.)
    count_(by) %>% # count claims
    left_join(denom, by=by) %>% # join n_bene
    mutate(cpkp=n*1000/n_bene) %>%
    select_(by, 'cpkp')
}

multicol_sum = function(x, new_name, target_columns) {
  dummy_vars = sapply(1:length(target_columns), function(i) sprintf('x%i', i))
  dummy_vars_string = paste0(dummy_vars, collapse=',')
  dummy_vars_list = lapply(target_columns, as.name) %>% setNames(dummy_vars)
  f = as.formula(paste0(c('~sum(', dummy_vars_string, ')'), collapse=''))

  rowwise(x) %>%
    mutate_(new_attribute=interp(f, .values=dummy_vars_list)) %>%
    rename_(.dots=setNames('new_attribute', new_name)) %>%
    ungroup()
}

@@

multivariate_formula = paste(c('n_claims ~ age_over_66 + is_female + ', conditions), collapse=' + ') %>% formula
start = rep(1.0, 3 + length(conditions))

all_condition_model = function(usage) {
  usage %>%
    mutate(age_over_66 = age - 66) %>%
    glm(multivariate_formula, data=., family=quasipoisson(link='identity'), start=start) %>%
    tidy()
}
```

```{r load_db, cache=FALSE}
regions = read_tsv('../../db/census-regions/census-regions.tsv') %>%
  select(state, region)
condition_names = read_tsv('../chronic-conditions/names.txt',
                           col_names=c('condition', 'condition_name'))
```

```{r analyze_function}
age_group_f = function(x) bounds_group(x, c(65, 70, 75, 80, 85, 90, 95, 100, 116))
claims_group_f = function(x) bounds_group(x, c(0, 1, 2, 5, 10, 20, Inf))

analyze = function(this_year) {
  # bene file is already filtered for plan coverage months and good geographies
  bene = read_tsv(sprintf('../bene_%i.tsv', this_year)) %>%
    filter(between(age, 66, 116), hmo_months==0) %>%
    mutate(is_female=sex=='female', is_white=race=='white') %>%
    left_join(regions, by='state')

  assert_has_no_duplicates(bene$bene_id)
  total_n_bene = nrow(bene)

  # keep only PDEs for known beneficiaries (semi join)
  # convert all amox/clav to just amox
  pde = read_tsv(sprintf('../pde_%i.tsv', this_year)) %>%
    semi_join(bene, by='bene_id') #%>%
    #mutate(antibiotic=if_else(antibiotic=='amoxicillin/clavulanate', 'amoxicillin', antibiotic))

  total_n_pde = nrow(pde)

  totals = data_frame(key=c('n_bene', 'n_pde'),
                      value=c(total_n_bene, total_n_pde),
                      year=this_year)

  # claims per 1,000 people by abx class
  claims_by_class = pde %>%
    count(antibiotic_class) %>%
    mutate(cpkp=n*1000/total_n_bene) %>%
    select(antibiotic_class, cpkp) %>%
    mutate(year=this_year)

  # the top few inidividual antibiotics
  top_abx = pde %>%
    count(antibiotic) %>%
    mutate(cpkp=n*1000/total_n_bene) %>%
    select(antibiotic, cpkp) %>% 
    mutate(year=this_year)

  # from here on, the denominators are taken from the beneficiary data grouped
  # in the same way as the consumption data, so we can use the summarize_by function
  claims_by_sex = summarize_by(bene, pde, 'sex') %>%
    mutate(year=this_year)

  # then by age group
  # (the input to summarize_by is modified, since we want to group by age *group*,
  # not just by raw age)
  claims_by_age = summarize_by(
      bene %>% mutate(age_group=age_group_f(age)),
      pde, 'age_group'
    ) %>%
    mutate(year=this_year)

  # then by census region
  claims_by_region = summarize_by(bene, pde, 'region') %>%
    mutate(year=this_year)

  # characteristics by consumption
  cc = read_tsv(sprintf('../cc_%i.tsv', this_year)) %>%
    select(-ALZH)

  usage = pde %>% count(bene_id) %>% rename(n_claims=n) %>%
    right_join(bene, by='bene_id') %>%
    replace_na(list(n_claims=0)) %>%
    left_join(cc, by='bene_id') %>%
    multicol_sum('n_cc', conditions)

  chars_by_consumption = usage %>%
    mutate(claims_group=claims_group_f(n_claims)) %>%
    group_by(claims_group) %>%
    summarize(`no. beneficiaries`=n(),
              `mean age`=mean(age),
              `% female`=mean(is_female) * 100,
              `mean no. chronic conditions`=mean(n_cc)) %>%
    mutate(year=this_year)

  condition_model = all_condition_model(usage) %>%
    mutate(year=this_year)

  list(totals=totals,
       claims_by_class=claims_by_class,
       top_abx=top_abx,
       claims_by_sex=claims_by_sex,
       claims_by_age=claims_by_age,
       claims_by_region=claims_by_region,
       chars_by_consumption=chars_by_consumption,
       condition_model=condition_model)
}

@@

totals = data_by_field('totals') %>%
  spread(key, value) %>%
  mutate(cpkp=n_pde*1000/n_bene) %>%
  mutate_at(vars(n_bene, n_pde), function(x) prettyNum(x, big.mark=',')) %>%
  rename(`number of beneficiaries`=n_bene, `number of antibiotic PDEs`=n_pde, `CPKP`=cpkp) %>%
  arrange(year) %>%
  kable

@@

top_classes = claims_by_class %>%
  filter(year==2011) %>%
  arrange(desc(cpkp)) %>%
  head(9) %$%
  antibiotic_class %>%
  c('other')

claims_by_class %<>%
  mutate(antibiotic_class=if_else(antibiotic_class %in% top_classes,
                                  antibiotic_class,
                                  'other')) %>%
  group_by(year, antibiotic_class) %>%
  summarize(cpkp=sum(cpkp)) %>%
  spread(year, cpkp) %>%
  mutate(antibiotic_class=factor(antibiotic_class, levels=top_classes, ordered=TRUE)) %>%
  arrange(antibiotic_class) %>%
  kable

top_abx = data_by_field('top_abx') %>%
  spread(year, cpkp) %>%
  arrange(desc(`2011`)) %>%
  head(11) %>%
  kable

claims_by_sex = data_by_field('claims_by_sex') %>%
  filter(sex %in% c('male', 'female')) %>%
  spread(year, cpkp) %>%
  kable

claims_by_age = data_by_field('claims_by_age') %>%
  filter(!is.na(age_group)) %>%
  spread(year, cpkp) %>%
  kable

claims_by_region = data_by_field('claims_by_region') %>%
  filter(!is.na(region)) %>%
  spread(year, cpkp) %>%
  arrange(desc(`2011`)) %>%
  kable

chars_by_consumption = data_by_field('chars_by_consumption') %>%
  arrange(desc(year), desc(claims_group)) %>%
  kable

condition_model = data_by_field('condition_model') %>%
  mutate(condition=str_replace(term, 'TRUE$', '')) %>%
  left_join(condition_names, by='condition') %>%
  mutate(term=if_else(is.na(condition_name), condition, condition_name)) %>%
  select(term, estimate, year) %>%
  spread(year, estimate) %>%
  arrange(desc(`2011`)) %>%
  kable
```

@@

`r claims_by_age`

>>>>

summarize_by = function(usage, by) {
  group_by_(usage, by) %>%
  summarize(n_claims=sum(n_claims), n_bene=n()) %>%
  mutate(cpkp=n_claims*1000/n_bene) %>%
  select_(by, 'cpkp')
}
```

```{r load_data}
regions = read_tsv('../../db/census-regions/census-regions.tsv') %>%
  select(state, region)
condition_names = read_tsv('../chronic-conditions/names.txt',
                           col_names=c('condition', 'condition_name'))

@@

load_data = function(year) {
  usage = read_tsv(sprintf('usage_%i.tsv', year))
  pde = read_tsv(sprintf('pde_%i.tsv', year))

  list(usage=usage, pde=pde)
}
```

```{r single_condition_model}
single_condition_model = function(x, condition) {
  filter(x, age >= 67) %>%
  mutate(age_over_67 = age - 67) %>%
  rename_(.dots=setNames(condition, 'has_condition')) %>%
    glm(n_claims ~ age_over_67 + is_female + has_condition,
        family=quasipoisson(link='identity'), data=.) %>%
    tidy() %>%
    mutate(condition=condition)
}

single_condition_models = function(usage) {
  lapply(conditions, function(cond) single_condition_model(usage, cond)) %>%
    bind_rows()
}
```

```{r analyze_function}
age_group_f = function(x) bounds_group(x, c(66, 70, 75, 80, 85, 90, 95, 100))
claims_group_f = function(x) bounds_group(x, c(0, 1, 2, 5, 10, 20, Inf))

analyze = function(year) {
  dat = load_data(year)

  total_n_bene = nrow(dat$usage)
  total_n_pde = nrow(dat$pde)

  totals = data_frame(key=c('n_bene', 'n_pde'),
                      value=c(total_n_bene, total_n_pde))

  # claims per 1,000 people by abx class
  claims_by_class = dat$pde %>%
    count(antibiotic_class) %>%
    mutate(cpkp=n*1000/total_n_bene) %>%
    select(antibiotic_class, cpkp)

  # the top few inidividual antibiotics
  top_abx = dat$pde %>%
    count(antibiotic) %>%
    mutate(cpkp=n*1000/total_n_bene) %>%
    select(antibiotic, cpkp)

  # from here on, the denominators are taken from the beneficiary data grouped
  # in the same way as the consumption data, so we can use the summarize_by function
  claims_by_sex = summarize_by(dat$usage, 'is_female')

  # then by age group
  # (the input to summarize_by is modified, since we want to group by age *group*,
  # not just by raw age)
  claims_by_age = dat$usage %>%
    group_by(age) %>%
    summarize(n_bene=n(),
              mean_n_claims=mean(n_claims),
              sd_n_claims=sd(n_claims),
              mean_n_cc=mean(n_cc),
              sd_n_cc=sd(n_cc)) %>%
    mutate(sem_n_claims=sd_n_claims/sqrt(n_bene),
           sem_n_cc=sd_n_cc/sqrt(n_bene))

  claims_by_age_group = dat$usage %>%
    mutate(age_group=age_group_f(age)) %>%
    group_by(age_group) %>%
    summarize(n_claims=sum(n_claims), n_bene=n()) %>%
    mutate(cpkp=n_claims*1000/n_bene) %>%
    select(age_group, n_bene, cpkp)

  claims_by_region = summarize_by(dat$usage, 'region')

  chars_by_consumption = dat$usage %>%
    mutate(claims_group=claims_group_f(n_claims)) %>%
    group_by(claims_group) %>%
    summarize(`no. beneficiaries`=n(),
              `mean age`=mean(age),
              `% female`=mean(is_female) * 100,
              `mean no. chronic conditions`=mean(n_cc))

  condition_model = single_condition_models(dat$usage)

  list(totals=totals,
       claims_by_class=claims_by_class,
       top_abx=top_abx,
       claims_by_sex=claims_by_sex,
       claims_by_age=claims_by_age,
       claims_by_age_group=claims_by_age_group,
       claims_by_region=claims_by_region,
       chars_by_consumption=chars_by_consumption,
       condition_model=condition_model) %>%
    lapply(function(x) mutate(x, year=year))
}

@@

totals = data_by_field('totals') %>%
  spread(key, value) %>%
  mutate(cpkp=round(n_pde*1000/n_bene)) %>%
  mutate_at(vars(n_bene, n_pde), function(x) prettyNum(x, big.mark=',')) %>%
  rename(`number of beneficiaries`=n_bene, `number of antibiotic PDEs`=n_pde, `CPKP`=cpkp) %>%
  arrange(year) %T>%
  write_tsv('tables/table1.tsv') %>%
  kable

@@

top_classes = claims_by_class %>%
  filter(year==2011) %>%
  arrange(desc(cpkp)) %>%
  head(9) %$%
  antibiotic_class

class_levels = c(top_classes, 'other')

claims_by_class %<>%
  mutate(antibiotic_class=if_else(antibiotic_class %in% top_classes,
                                  antibiotic_class,
                                  'other')) %>%
  group_by(year, antibiotic_class) %>%
  summarize(cpkp=round(sum(cpkp))) %>%
  spread(year, cpkp) %>%
  mutate(antibiotic_class=factor(antibiotic_class, levels=class_levels, ordered=TRUE)) %>%
  arrange(antibiotic_class) %T>%
  write_tsv('tables/table2_class.tsv') %>%
  kable

top_abx = data_by_field('top_abx') %>%
  mutate(cpkp=round(cpkp)) %>%
  spread(year, cpkp) %>%
  arrange(desc(`2011`)) %>%
  head(11) %T>%
  write_tsv('tables/table2_drug.tsv') %>%
  kable

claims_by_sex = data_by_field('claims_by_sex') %>%
  mutate(cpkp=round(cpkp)) %>%
  spread(year, cpkp) %T>%
  write_tsv('tables/table2_sex.tsv') %>%
  kable

claims_by_age_group = data_by_field('claims_by_age_group') %>%
  mutate(cpkp=round(cpkp)) %>%
  select(age_group, year, cpkp) %>%
  spread(year, cpkp) %T>%
  write_tsv('tables/supptable1_age_group.tsv') %>%
  kable

bene_by_age_group = data_by_field('claims_by_age_group') %>%
  mutate(n_bene=prettyNum(n_bene, big.mark=',')) %>%
  select(age_group, year, n_bene) %>%
  spread(year, n_bene) %T>%
  write_tsv('tables/supptable1_age_group_denom.tsv') %>%
  kable

claims_by_age = data_by_field('claims_by_age')

claims_by_region = data_by_field('claims_by_region') %>%
  mutate(cpkp=round(cpkp)) %>%
  spread(year, cpkp) %>%
  arrange(desc(`2011`)) %T>%
  write_tsv('tables/table2_region.tsv') %>%
  kable

chars_by_consumption = data_by_field('chars_by_consumption') %>%
  mutate_at(vars(`mean age`, `% female`, `mean no. chronic conditions`), function(x) round(x, digits=2)) %>%
  arrange(year, claims_group) %>%
  kable

condition_model = data_by_field('condition_model') %>%
  left_join(condition_names, by='condition') %>%
  filter(year==2011, term != '(Intercept)') %>%
  mutate(estimate=round(estimate, digits=3)) %>%
  select(condition=condition_name, term, estimate) %>%
  spread(term, estimate) %>%
  arrange(desc(has_conditionTRUE)) %T>%
  write_tsv('tables/table3.tsv') %>%
  kable
```

@@

`r claims_by_age_group`

`r bene_by_age_group`

# Figure 1: Claims and conditions by age

```{r fig1_claims_by_age}
claims_by_age %>%
  filter(year==2011) %>%
  ggplot(aes(x=age,
             y=mean_n_claims,
             ymin=mean_n_claims-sem_n_claims,
             ymax=mean_n_claims+sem_n_claims)) +
  geom_point() +
  geom_errorbar() +
  theme_minimal() +
  scale_x_continuous(breaks=c(66, 70, 80, 90, 100))
```

```{r fig1_cc_by_age}
claims_by_age %>%
  filter(year==2011) %>%
  ggplot(aes(x=age,
             y=mean_n_cc,
             ymin=mean_n_cc-sem_n_cc,
             ymax=mean_n_cc+sem_n_cc)) +
  geom_point() +
  geom_errorbar() +
  theme_minimal() +
  scale_x_continuous(breaks=c(66, 70, 80, 90, 100))
```

### 2017-04-12 ms1/supp1.Rmd
updated

```{r single_condition}
summarize_condition = function(x, condition) {
rename_(x, .dots=setNames(condition, 'has_condition')) %>%
  glm(n_claims ~ age_over_66 + is_female + has_condition, family='poisson', data=.) %>%
  tidy() %>%
  mutate(condition=condition)
}

single_condition_models = lapply(conditions, function(cond) summarize_condition(usage, cond)) %>%
  bind_rows()
```

>>>>

```{r all_condition_model}
multivariate_formula = paste(c('n_claims ~ age_over_66 + is_female + ', conditions), collapse=' + ') %>% formula
start = rep(1.0, 3 + length(conditions))

all_condition_model = function(usage) {
  usage %>%
    mutate(age_over_66 = age - 66) %>%
    glm(multivariate_formula, data=., family=quasipoisson(link='identity'), start=start) %>%
    tidy()
}
```
